---
editor_options: 
  markdown: 
    wrap: 72
---

#########Sleep of Mammals############# 
#(a) Describe the dataset using appropriate plots/curves/charts,...

```{r}
data=read.csv(file.choose())
data

```

```{r}
t=data$brain_wt
plot(t,type='l',col=("Blue"))
curve(x^4,from=10,to=-10)
c=data$danger
pie(table(c))
barplot(table(c))

```

\#(b) Consider one of continuous attributes, and compute central and
variational measures.

```{r}
N=data$exposure
y=table(N)
y
mode=names(y)[which(y==max(y))]
c("mode=",mode)

```

```{r}

X=data$body_wt
m=mean(X)
c("mean=",m)


```

```{r}
med=median(X)
c("Median=",med)

```

```{r}
s=sd(X)
c("Standard deviation=",s)

```

#Data Normalization(Z-Score)

```{r}
Z=(X-m)/s
Z

```

```{r}
R=max(X)-min(X)
R

```

#Scaling

```{r}
Y=(X-min(X))/R
Y
dh=data.frame(X,Y,Z)
head(dh)
colnames(dh)=c('body_wt','scaled body_wt','Normalized body_wt')
head(dh)
```

\#(c) For a particular variable of the dataset, use Chebyshev's rule,
and propose one-sigma interval. Based on your proposed interval, specify
the outliers if any.

```{r}
L=mean(X)-sd(X); U=mean(X)+sd(X)
c("Lower_bound=",L,"Upper_bound=",U)
out=c()
for(j in X){
  if(j>U|j<L){o=j}else{o=NULL}
  out=c(out,o)
}
c('Outliers=',out)
```

\#(d) Explain how the box-plot technique can be used to detect outliers.
Apply this technique for one attribute of the dataset

```{r}
Q1=quantile(X,0.25);Q3=quantile(X,0.75)
IQR=Q3-Q1
IQR
LW=Q1-1.5*IQR(X);UW=Q3+1.5*IQR(X)
out=c()
for(i in X){
  if(i>UW|i<LW){o=i}else{o=NULL}
  out=c(out,o)
}
c("Lower_bound=",LW,"Upper_bound=",UW)
c("Outliers=",out)
```

#filling the NULL space

```{r}
N=data$exposure
y=table(N)
y
mode=names(y)[which(y==max(y))]
c("mode=",mode)
data$gestation[is.na(data$gestation)]=mode
data$life_span[is.na(data$life_span)]=mode
X=data$body_wt
m=mean(X)
c("mean=",m)
data$non_dreaming[is.na(data$non_dreaming)]=m
data$dreaming[is.na(data$dreaming)]=m
data$total_sleep[is.na(data$total_sleep)]=m
data
```

########End of the First Question###########
#    2a)Select four variables of the dataset, and propose an appropriate probability model to quantify uncertainty of each variable. 
#b)	 For each model in part (a), estimate the parameters of model.     
#c)	Express the way in which each model can be used for the predictive analytics, then find the prediction for each attribute.                              

#Reading the dataset
```{r}
data=read.csv(file.choose())
data

```

```{r}
X=data$charges
N=data$children
J=data$bmi
X=data$region
```


#Exponentional
```{r}
X=data$charges
xbar=mean(X);xbar
alpha=(xbar^2)/var(X);alpha
lambda=xbar/var(X);lambda
rexp(X,lambda)
```
```{r}
L=mu-3*sigma; U=mu+3*sigma
L;U
j=seq(L,U,10);j
pdf=dexp(j,lambda)
pdf
```

```{r}
plot(j,pdf,type='b',col='green',main='Exponantional',xlab='changed',ylab='prob density function')
```

```{r}
cdf=pexp(j,lambda)
cdf

```

```{r}
plot(j,cdf,type='b',col='red',main='Exponantional',xlab='Charges',ylab='cumulative distribution function')
```

```{r}
Predection=sum(dnorm(j,lambda))
Predection
```

#Binomial Distribution
```{r}
N=data$children
n=max(N);n
p=mean(N)/n
p
```

```{r}
hyperparameters=rbinom(N,n,p)
hyperparameters
```


```{r}
# P(X=3)
dbinom(3,n,p)
# p(X<=5)
pbinom(5,n,p)
j=0:n;j
#pmf
pmf=dbinom(j,n,p)
c(pmf)
cdf=pbinom(j,n,p)
c(cdf)
```
```{r}
plot(j,cdf, type='b', col='skyblue')
lines(j,pmf, type='b', col='green')
```

```{r}
#  E(X)
E=n*p
E
```

```{r}
Predection=sum(dbinom(j,n,p))
Predection
```
#  Normal distribution
```{r}
J=data$bmi
#parameter estimation
mu=mean(J); sigma=sd(J)
mu; sigma
```

```{r}
rnorm(J,mu,sigma)
```

```{r}
L=mu-2*sigma; U=mu+2*sigma
L;U
```

```{r}
j=seq(L,U,1);j
```

```{r}
pdf=dnorm(j,mu,sigma)
pdf

```

```{r}
plot(j,pdf,type='b',col='green',main='normal pdf of bmi',xlab='bmi',ylab='prob density function')
```

```{r}
cdf=pnorm(j,mu,sigma)
cdf

```

```{r}
plot(j,cdf,type='b',col='red',main='normal pdf of price',xlab='bmi',ylab='cumulative distribution function')
```

```{r}
Predection=sum(dnorm(j,mu,sigma))
Predection

```
# multinomil distribution 
```{r}
X=data$region
f=table(X); f
p=f/sum(f);p 
```

```{r}
mode=names(f)[which(f==max(f))]; mode
c("prediction of region using multinomil distribution :", mode)
```




y=table(N)
y
mode=names(y)[which(y==max(y))]
mode
data$gestation[is.na(data$gestation)]=mode
data$life_span[is.na(data$life_span)]=mode
X=data$body_wt
m=mean(X)
m
data$non_dreaming[is.na(data$non_dreaming)]=m
data$dreaming[is.na(data$dreaming)]=m
data$total_sleep[is.na(data$total_sleep)]=m
data
var1=data$non_dreaming
var2=data$total_sleep
var3=data$gestation
var4=data$exposure